% 1. 替换为华科博士学位论文模板
\documentclass[type=doctor, anonymous=true]{hustthesis} 

%========== 论文信息配置 (hustsetup) ==========
\hustsetup { 
  info =
  {
    title = { 基于里德堡原子阵列的量子多体动力学研究 },
    title* = { Research on Quantum Many-Body Dynamics Based on Rydberg Atom Arrays }, % 需补充英文标题
    degree = { academic }, 
    degree* = { Physics }, 
    author = { 项德生 },
    author* = { XIANG Desheng },
    student-id = { D202080053 },
    clc = {  }, % 中图分类号（示例）
    major = { 物理学 },
    major* = { Physics },
    supervisor = { 李霖\quad{} 教授 },
    supervisor* = { Prof.~ LI Lin },
    % date = { 2025-02-28 }, % 或直接删除此行由系统根据编译日期生成
    % committee = { 张三 & 教授 & 华中科技大学 , 李四 & 教授 & 武汉大学 }, % 答辩时填写
  },
}

%========== 宏包引入 ==========
\usepackage{subfiles}
\usepackage{amsmath,amssymb}
\usepackage{graphicx}
\usepackage{rotating}
\usepackage{dcolumn}
\usepackage{bm}
\usepackage{siunitx}
% \usepackage[backend=biber,style=numeric]{biblatex} % 模板和原文件都使用 biblatex
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{makecell}
\usepackage{physics}
\usepackage{amsfonts}
\usepackage{todonotes}

% 引入你的参考文献库
\addbibresource{preamble/references.bib} 

%========== 自定义命令 ==========
\newcommand{\LL}[1]{\textcolor[RGB]{153,51,250}{#1}}
\newcommand{\XDS}[1]{\textcolor[RGB]{250,250,250}{#1}}
\renewcommand\theadfont{\bfseries}

\begin{document}

% 自动生成符合学校规范的各类封面页和标题页
\maketitle 

\frontmatter

%========== 摘要部分 ==========
% 使用模板专用的 abstract 环境
\begin{abstract}
强关联量子多体系统在远离平衡态时的动力学演化，是涵盖凝聚态物理、量子信息科学至高能规范场论的核心前沿议题。受限于希尔伯特空间的指数级膨胀与符号问题，经典计算难以对复杂多体动力学进行精确求解。基于光镊阵列的中性原子可编程量子模拟器，凭借其单粒子级别的空间排布自由度与可调控的强里德堡相互作用，为突破上述算力瓶颈提供了理想的物理平台。本文以探索受限多体希尔伯特空间中的非平衡物理为目标，从全栈式实验平台的硬件构筑出发，系统性地研究了量子多体疤痕态、量子信息加扰与回流，以及格点规范场论中的实时散射动力学。

在实验方法与平台构筑方面，本文自主设计并搭建了一套基于铷-87原子的超高真空双腔量子操控系统。通过结合空间光调制器（SLM）与声光偏转器（AOD），实现了大规模无缺陷二维单原子阵列的确定性重排。针对高保真度操控与探测的需求，开发了基于极化梯度机制的无损单原子成像技术；利用自由空间相位调制结合体光栅，构建了远失谐受激拉曼光路以实现基态相干操控；并通过探究玻璃腔内壁的动态电荷屏蔽物理机制，创新性地提出了瞬态多脉冲电场电离方案，将里德堡态的探测保真度提升至物理极限水平。

在量子多体动力学演化方面，针对全局绝热演化在系统尺寸扩大时面临的能隙缩小瓶颈，本文提出基于局域光频移的单步确定性初态制备方案，在多体阵列中实现了高纯度量子多体疤痕态的制备。以此为基准，利用里德堡系统的粒子-空穴对称性结合全局微波驱动，本文在实验上成功执行了多体哈密顿量的时间反演，首次测得了非时序关联函数（OTOC）的时空动力学，揭示了量子信息在受限多体空间中沿光锥传播并发生周期性塌缩与复苏的物理图景。为打破特定微扰算符的观测盲区，本文进一步引入局域耗散与电磁诱导透明（EIT）屏蔽技术，在强相互作用背景下实现了单地址分辨的受限量子态层析。通过提取 Holevo 信息与迹距离，定量确证了系统存在强烈的非马尔可夫量子信息回流机制，并从理论与实验上严格排除了该微观信息复苏作为波函数振荡平庸副产物的可能。

在跨学科物理探索方面，本文将里德堡阻塞的局域动力学约束映射为（1+1）维格点规范场论（LGT）中的高斯定律。借助纳秒级时空分辨率的哈密顿量工程，本文确定性地制备了类介子多体激发，并动态精确调控交错费米子质量与有效弦张力参量。实验不仅观测到了由量子淬火驱动的禁闭-解禁闭相变及弦破碎特征，还在解禁闭相中实时追踪了类介子的准弹性散射与干涉动力学。尤为突破的是，通过在碰撞波包重叠的最激烈时刻对体系施加瞬态禁闭淬火，本文成功截获了高度纠缠的散射中间态，观测到了强相互作用多体演化中的动力学冻结现象，在微观层面上与高能物理中重离子碰撞的演化图像形成了深刻的映射。

综上所述，本文系统性地完成了从底层量子模拟器构筑到复杂多体物理探索的研究工作。本研究不仅在实验硬件与高保真度操控技术上提供了可靠的工程积累，其所观测到的受限多体动力学现象与展现的时空控制能力，也为进一步探究各态历经性破缺机制、微观量子信息输运界限以及高能格点规范场动力学提供了坚实的实验基础与详实的数据参考。
  \keywords 量子模拟，里德堡原子，非平衡动力学，量子多体疤痕，非时序关联函数，格点规范场论
\end{abstract}

\begin{abstract*}
The non-equilibrium dynamical evolution of strongly correlated quantum many-body systems represents a core frontier bridging condensed matter physics, quantum information science, and high-energy gauge field theories. Limited by the exponential expansion of the Hilbert space and the severe sign problem, classical computation struggles to exactingly simulate complex many-body dynamics. Programmable quantum simulators based on neutral atom optical tweezer arrays, leveraging their single-particle-level spatial configuration degrees of freedom and microsecond-scale strong Rydberg interactions, provide an ideal physical platform to overcome the aforementioned computational bottlenecks. Aiming to explore non-equilibrium physics within constrained many-body Hilbert spaces, this thesis systematically investigates quantum many-body scars, quantum information scrambling and backflow, as well as real-time scattering dynamics in lattice gauge theories, starting from the hardware construction of a full-stack experimental platform.

Regarding experimental methods and platform construction, this thesis independently designs and builds an ultra-high vacuum dual-chamber quantum manipulation system based on Rubidium-87 atoms. By integrating spatial light modulators (SLM) and acousto-optic deflectors (AOD), the deterministic rearrangement of large-scale, defect-free two-dimensional single-atom arrays is achieved. To meet the demands for high-fidelity manipulation and detection, a non-destructive single-atom imaging technique based on the polarization gradient cooling mechanism is developed; a far-detuned stimulated Raman optical setup is constructed using free-space phase modulation combined with a volume Bragg grating to realize coherent ground-state manipulation; furthermore, by investigating the dynamic charge shielding mechanism on the inner wall of the glass cell, an innovative transient multi-pulse electric field ionization scheme is proposed, pushing the detection fidelity of Rydberg states to the physical limit.

In the realm of quantum many-body dynamical evolution, addressing the shrinking energy gap bottleneck faced by global adiabatic evolution when scaling up system sizes, this thesis proposes a single-step deterministic initial state preparation scheme based on local optical light shifts, achieving a record-high purity in the preparation of quantum many-body scar states in a many-body array. Establishing this as a benchmark, and exploiting the particle-hole symmetry of the Rydberg system combined with global microwave driving, this thesis experimentally executes the time reversal of the many-body Hamiltonian. For the first time, the spatiotemporal dynamics of the out-of-time-ordered correlator (OTOC) is measured, revealing a physical picture where quantum information propagates along a light cone and undergoes periodic collapse and revival within the constrained many-body space. To overcome the observation blind spots of specific perturbation operators, this thesis further introduces local dissipation and electromagnetically induced transparency (EIT) shielding techniques, achieving single-site-resolved constrained quantum state tomography under strong interactions. By extracting Holevo information and trace distance, the thesis quantitatively confirms the existence of a strong non-Markovian quantum information backflow mechanism in the system, rigorously ruling out theoretically and experimentally the possibility that this microscopic information revival is merely a trivial byproduct of wavefunction oscillations.

In terms of cross-disciplinary physics exploration, this thesis maps the local dynamical constraints of the Rydberg blockade to Gauss's law in a (1+1)-dimensional lattice gauge theory (LGT). Empowered by Hamiltonian engineering with nanosecond spatiotemporal resolution, this thesis deterministically prepares meson-like many-body excitations and dynamically tunes the staggered fermion mass and effective string tension parameters. The experiment not only observes the confinement-deconfinement transitions and string fragmentation features driven by quantum quenches, but also real-time tracks the quasi-elastic scattering and interference dynamics of mesons in the deconfined phase. More breakthroughingly, by applying a transient confinement quench exactly at the most intense moment of colliding wavepacket overlap, this thesis successfully intercepts the highly entangled intermediate scattering state, observing a dynamical freeze-out phenomenon in strongly interacting many-body evolution, which forms a profound microscopic mapping to the evolution picture of heavy-ion collisions in high-energy physics.

In conclusion, this thesis systematically accomplishes a comprehensive research endeavor spanning from the construction of a foundational quantum simulator to the exploration of complex many-body physics. This research not only provides reliable engineering accumulations in experimental hardware and high-fidelity manipulation techniques, but the observed constrained many-body dynamical phenomena and demonstrated spatiotemporal control capabilities also provide a solid experimental foundation and detailed data reference for further exploring ergodicity breaking mechanisms, microscopic quantum information transport bounds, and high-energy lattice gauge field dynamics.

\textbf{Keywords:} Quantum simulation, Rydberg atoms, Non-equilibrium dynamics, Quantum many-body scars, Out-of-time-ordered correlator, Lattice gauge theory
\end{abstract*}

% 生成目录
\tableofcontents

% 添加插图清单
\cleardoublepage
\listoffigures

% 添加表格清单
\cleardoublepage
\listoftables

\mainmatter

%========== 正文章节包含 ==========
% 保留原有的 subfile 结构
\subfile{chapters/chapter_01/01-introduction}
\subfile{chapters/chapter_02/02-theory}
\subfile{chapters/chapter_03/03-setup}
\subfile{chapters/chapter_04/04-Rydberg_control}
\subfile{chapters/chapter_05/05-quantum_scar}
\subfile{chapters/chapter_06/06-OTOC}
\subfile{chapters/chapter_07/07-LGT}
\subfile{chapters/chapter_08/08-Outlook}

\backmatter

%========== 致谢部分 ==========
% 替换为模板专用的 acknowledgements 环境
\begin{acknowledgements} 
  感谢所有老师与同学。
\end{acknowledgements}

%========== 参考文献 ==========
\printbibliography

\appendix

\chapter{附录}
% 可在此处添加附录内容或通过 \subfile 引入

\end{document}